Inferring generation-interval distributions from contact tracing data: \\ \emph{in prep, PRSB}

### 2020-05-20 doc/response.tex

\rev{
The paper focuses on estimating the generating-interval distribution. There are
several generating-intervals and distributions mentioned in this paper, such as
$g(\tau)$; $b_t(\tau)$; $c_t(\tau)$; $g_eff(\tau)$; etc. To the end, it seems that the primary focus is the
intrinsic distribution $g(\tau)$; and relating this distribution with two other intrinsic
parameters: the reproduction number $\Ro$ and the Malthusian number $r$
}

The intrinsic distribution is not our primary focus. 
As we show in Figure 4, the intrinsic distribution overesitmates $\Rini$ as it does not account for spatial effects.
We tried to make clear that our goal is the initial forward distribution.

\rev{
The authors have well demonstrated different sources of sampling bias
arising from contact tracing data, more noticeably, (i) length-bias due
to right-truncation because of retrospective measurement together with
different infection times during the course of the outbreak; (ii) due to the
network structure (e.g. the degree distribution and the degree excess (Line
408). Both sources, the right-truncation and the egocentric point of view
(Line 174), tend to give shorter on average observed intervals. These are
correct and important.
}

\rev{
Any notion of bias is in the context against some "truth". The authors
proposed, and demonstrated by simulation, ways to adjust for the bias in
order to estimate this "truth". However, the authors equate the "true"
distribution to the intrinsic distribution $g(\tau)$ as if "obvious". It is not
obvious and this needs to be proven.
}

We had never equated the true distribution to the intrinsic distribution.
We make clear that we do not want the intrinsic distribution.

\rev{
The intrinsic distribution $g(\tau)$ is an equilibrium distribution that
involves the equilibrium distribution of the infectious period (in SEIS,
SEIR, SEIRS settings). It is a renewal process concept and is only
valid when the system is under the equilibrium condition. It is not
obvious how it is related to measurements of "generation intervals"
when the system is not at equilibrium.
}

\rev{
 Under suitable assumptions on homogeneous mixing, the intrinsic
distribution $g(\tau)$ measures the infectiousness within an infected individual without involving another individual. It is not obvious how it
is related to measures involving two individuals, no matter whether
they are prospectively or retrospectively measured.
}

\rev{
Potentially observable events do not take place at the exact moment
when transmission occurs.
}

\rev{
I rate this paper very high in scientific importance and in general interest.
In the last 20 years, especially since SARS outbreak in 2003, there has
been a huge volumn of related work, more than those referenced in this
paper. These "intervals" have been widely cited, measured and applied to
outbreak investigations, especially during the early stage. However, there
is a lot of ambiguity. There are occasions that the same terminology is
associated with two different definitions, whereas there are other occasions
that the same definition is assigned to different terminologies by different
researchers.
}

\rev{
I suggest major revision of the paper and hope my comments are
constructive for the revision(s) for better clarification. It will be a
great contribution to this field.
}

\rev{
The immediate objective of this paper is to construct some "unbiased" estimate of the generation interval distribution (presumably, the intrinsic distribution $g(\tau)$). Its important goes beyond this immediate objective, that is, to attempt using quantities (assuming observable, at least conceptually) measured during a the very dynamic
phase of the system to make inferences on these intrinsic quantities
$\Ro$ and $r$: I feel the authors have fallen short of achieving these objectives. However, I would like to share some of my opinions and ideas
and hope that, after revision(s), some of these goals will be achieved.
}

\rev{
I also recommend, for the authors to consider, breaking this paper
into two papers: one on mathematical theory and models to better
clarify concepts which I believe are still ambiguous in currently published literature; and another one on statistical models and methods
to link data to models.
}
